Making Sequential Programs Deterministic

### 2019-03-26 design.tex
poly resolution table
poly resolution
type refinement

Valid locations for \codeid{@PolyDet("up")} and  \codeid{@PolyDet("down")}:

$\infer{\vdash : \kappa \  \beta}{isParam(\beta) / isReturn(\beta) & \kappa : @PolyDet("up") / @PolyDet("down")}$

Resolution of \<@PolyDet("up")>:\todo{typesetting}

$\infer{\Gamma \vdash x : @NonDet \ \beta_a}{\Gamma \vdash x : \kappa_a \ \beta_a &  \kappa_a : @OrderNonDet/@NonDet & \kappa_p : @PolyDet("up") & \kappa_p \ \beta_p : declaration \ type & \kappa_a \ \beta_a : resolved \ type}$

Resolution of \<@PolyDet("down")>:

$\infer{\Gamma \vdash x : @Det \  \beta_a}{\Gamma \vdash x :  \kappa_a \beta_a & \kappa_a : @OrderNonDet/@Det & \kappa_p : @PolyDet("down") & \kappa_p \ \beta_p : declaration \ type & \kappa_a \ \beta_a : resolved \ type}$

@@

@@

The determinism checker has additional type refinement rule for the sorting methods in collections.

The checker type refines a receiver of type \codeid{@OrderNonDet List} to a \codeid{@Det List} if all the
type argument of this \codeid{List} have the type annotation \codeid{@Det}. Similar rules apply for \codeid{Arrays.sort()} and
\codeid{Collections.sort()}.
Under this if \codeid{sort()} is called on a receiver of type \codeid{@OrderNonDet List<@Det Set<@Det String>>}, it will
be type refined to \codeid{@Det List<@Det Set<@Det String>>}. However a receiver of type \codeid{@OrderNonDet List<@OrderNonDet Set<@Det String>>} will not be type refined.

>>>>

% Not true: We allow Poly(up) and down at every location where Poly is allowed.
%Valid locations for \codeid{@PolyDet("up")} and  \codeid{@PolyDet("down")}:
%$\infer{\vdash : \kappa \  \beta}{isParam(\beta) / isReturn(\beta) & \kappa : @PolyDet("up") / @PolyDet("down")}$
%
%Resolution of \<@PolyDet("up")>:\todo{typesetting}
%
%$\infer{\Gamma \vdash x : @NonDet \ \beta_a}{\Gamma \vdash x : \kappa_a \ \beta_a &  \kappa_a : @OrderNonDet/@NonDet & \kappa_p : @PolyDet("up") & \kappa_p \ \beta_p : declaration \ type & \kappa_a \ \beta_a : resolved \ type}$
%
%Resolution of \<@PolyDet("down")>:
%
%$\infer{\Gamma \vdash x : @Det \  \beta_a}{\Gamma \vdash x :  \kappa_a \beta_a & \kappa_a : @OrderNonDet/@Det & \kappa_p : @PolyDet("down") & \kappa_p \ \beta_p : declaration \ type & \kappa_a \ \beta_a : resolved \ type}$

@@

\begin{figure}
    $\infer[\sc{\mathit{POLY RESOLUTION}}]{\pi_r \ \beta_r method(\overline{\pi_i \ \beta_i})\{\ \}}{\forall K \in \{\mathit{NonDet, OrderNonDet, Det\}} : K \ \beta_r method(\overline{K \ \beta_i})\{\ \} & \pi_r, \pi_i \in \{\mathit{PolyDet}, \mathit{PolyDet}\uparrow, \mathit{PolyDet}\downarrow    \}}$
\end{figure}

\begin{table}[]
    \begin{tabular}{|l|l|l|}
        \hline
        \textbf{\<PolyDet>} & \textbf{\<PolyDet$\uparrow$>} & \textbf{\<PolyDet$\downarrow$>} \\ \hline
        \<NonDet> & \<NonDet> &  \<NonDet>\\ \hline
        \<OrderNonDet> & \<NonDet> &  \<Det>\\ \hline
        \<Det> & \<Det> &  \<Det>\\ \hline
    \end{tabular}
\caption{Polymorphic resolution}
\label{tab-poly-resolutions}
\end{table}

@@

The determinism checker has additional type refinement rule for the sorting methods as shown in Figure~\ref{fig-sorting}.

\begin{figure}
    $\infer[\sc{\mathit{LIST \  SORT}}]{\Gamma \vdash x : \mathit{Det \ List}\langle\mathit{Det} \ \beta\rangle}{\Gamma \vdash x : \mathit{OrderNonDet \ List}\langle\mathit{Det} \ \beta\rangle & \mathit{x.sort()}}$

    $\infer[\sc{\mathit{COLLECTIONS \  SORT}}]{\Gamma \vdash x : \mathit{Det \ List}\langle\mathit{Det} \ \beta\rangle}{\Gamma \vdash x : \mathit{OrderNonDet \ List}\langle\mathit{Det} \ \beta\rangle & \mathit{Collections.sort.sort(x)}}$

    $\infer[\sc{\mathit{ARRAYS \  SORT}}]{\Gamma \vdash x : \mathit{Det \ \beta \ Det[]} }{\Gamma \vdash x : \mathit{Det \ \beta \ OrderNonDet[]} & \mathit{Arrays.sort( \ x)}}$

    \caption{Sorting type refinement rules}
    \label{fig-sorting}
\end{figure}

The checker type refines a receiver of type \codeid{@OrderNonDet List} to a \codeid{@Det List} if all the
type argument of this \codeid{List} have the type annotation \codeid{@Det} (Rule LIST SORT in Figure~\ref{fig-sorting}). Similar rules apply for \codeid{Arrays.sort()} and \codeid{Collections.sort()}.
As per these rules, if \codeid{sort()} is called on a receiver of type \codeid{@OrderNonDet List<@Det Set<@Det String>>}, it will
be type refined to \codeid{@Det List<@Det Set<@Det String>>}. However a receiver of type \codeid{@OrderNonDet List<@OrderNonDet Set<@Det String>>} will not be type refined.
